---
title: 'Project 3: Census Data'
author: "Devin Luu, Daniel Stroik, Michael Grube, Mark Dillemuth"
resource_files:
- .Renviron
output:
  html_notebook:
    code_folding: hide
    toc: yes
    toc_depth: 4
    toc_float: yes
runtime: shiny
---

```{r setup, include=FALSE}
library(tidyverse)
library(data.world)
library(shiny)
library(gridExtra) #grid.arrange()

knitr::opts_chunk$set(echo = TRUE)
```
  
## **R Session Info**  

```{r}
sessionInfo()
```

## **Github Link** 
https://github.com/devtl/f17-edv-project3

## **Introduction** 
Fall 2017 Data Visualization Project. We want to explore tobacco usage among highschool teenagers using data from the U.S. Census Bureau.

Dataset is stored on https://data.world/devtl/f-17-edv-project-3.

## **Getting data from data.world** 
We will be using the Fatal Police Shootings data and some census data from the U.S. Census Bureau from [data.world](https://data.world/devtl/fatalpoliceshootings/)

```{r}
project <- "https://data.world/devtl/f-17-edv-project-3/"
data.world::set_config(cfg_env("DW_API"))
df <- data.world::query(data.world::qry_sql("select smoking.surveyyear as year,
    smoking.statecode as state,
    poverty.geoid,
    ((poverty.b17001_008 + poverty.b17001_009) + (poverty.b17001_022 + poverty.b17001_023))/poverty.b17001_002 as belowpoverty_highschool,
    (poverty.b17001_008 + poverty.b17001_009)/poverty.b17001_002 as belowpoverty_highschool_m,
    (poverty.b17001_022 + poverty.b17001_023)/poverty.b17001_002 as belowpoverty_highschool_f,
    smoking.EverTryCigT as ever_tried,
    smoking.EverTryCigF as evertried_f,
    smoking.EverTryCigM as evertried_m,
    smoking.CigBefore13T as before13,
    smoking.CigBefore13F as before13_f,
    smoking.CigBefore13M as before13_m,
    smoking.CurrentCigT as currentsmoker,
    smoking.CurrentCigF as currentsmoker_f,
    smoking.CurrentCigM as currentsmoker_m,
    smoking.FrequentCigT as freqsmoker,
    smoking.FrequentCigF as freqsmoker_f,
    smoking.FrequentCigM as freqsmoker_m,
    smoking.TenCigDayT as 10aday,
    smoking.TenCigDayF as 10aday_f,
    smoking.TenCigDayM as 10aday_m,
    smoking.NotTryQuitT as notquit,
    smoking.NotTryQuitF as notquit_f,
    smoking.NotTryQuitM as notquit_m,
    smoking.CigAtSchoolT as atschool,
    smoking.CigAtSchoolF as atschool_f,
    smoking.CigAtSchoolM as atschool_m,
    smoking.BuyCigAtStoreT as buyscigs,
    smoking.BuyCigAtStoreF as buyscigs_f,
    smoking.BuyCigAtStoreM as buyscigs_m,
    smoking.EverCigDailyT as atleast1_daily,
    smoking.EverCigDailyF as atleast1_daily_f,
    smoking.EverCigDailyM as atleast1_daily_m,
    smoking.CigAll30DaysT as 30days,
    smoking.CigAll30DaysT as 30days_f,
    smoking.CigAll30DaysT as 30days_m,
    smoking.CurrentSmokelessT as smokeless,
    smoking.CurrentSmokelessT as smokeless_f,
    smoking.CurrentSmokelessT as smokeless_m,
    smoking.CurrentCigarT as cigar,
    smoking.CurrentCigarF as cigar_f,
    smoking.CurrentCigarM as cigar_m,
    smoking.CurrentAnyTobaccoT as anytobacco,
    smoking.CurrentAnyTobaccoF as anytobacco_f,
    smoking.CurrentAnyTobaccoM as anytobacco_m
    
from health.`yrbs-state-tobacco-variables`.yrbs_state_tobacco_variables_2013_v2 as smoking
join uscensusbureau.`acs-2015-5-e-poverty`.usa_all_states as poverty 
    on smoking.statecode = poverty.state"), dataset = project)
summary(df)
```


##Scatterplots{.tabset .tabset-fade}

Various Tobacco Usages vs. Highschool Poverty Levels

###Any type of Tobacco Usage
```{r}
df1 = df %>% dplyr::select(belowpoverty_highschool, anytobacco, atleast1_daily, atschool, before13, currentsmoker, freqsmoker, state)

renderPlot({
  pl1 = ggplot(df1, mapping = aes(belowpoverty_highschool, anytobacco, color = state)) + geom_point(size = 1.5) + theme_minimal()
  
 pl1 
})
```

###Daily Smokers
```{r}
renderPlot({
  pl2 = ggplot(df1, mapping = aes(belowpoverty_highschool, atleast1_daily, color = state)) + geom_point(size = 1.5) + theme_minimal()
  
  pl2
  
})
```


###Smokes at School
```{r}
renderPlot({
  pl3 = ggplot(df1, mapping = aes(belowpoverty_highschool, atschool, color = state)) + geom_point(size = 1.5) + theme_minimal()
  
  pl3 
})
```


###Smoked before 13 years old
```{r}
renderPlot({
  pl4 = ggplot(df1, mapping = aes(belowpoverty_highschool, before13, color = state)) + geom_point(size = 1.5) + theme_minimal()
  
  pl4

})
```


###Current Smoker
```{r}
renderPlot({
  pl5 = ggplot(df1, mapping = aes(belowpoverty_highschool, currentsmoker, color = state)) + geom_point(size = 1.5) + theme_minimal()
  
  pl5

})
```

###Frequent Smoker
```{r}
renderPlot({
  pl6 = ggplot(df1, mapping = aes(belowpoverty_highschool, freqsmoker, color = state)) + geom_point(size = 1.5) + theme_minimal()
  
  pl6
})
```

###Findings

We see that there is no evidence that poverty affects any smoking factors among highschool teens.


















