---
title: 'Project 3: Census Data'
author: "Devin Luu, Daniel Stroik, Michael Grube, Mark Dillemuth"
resource_files:
- .Renviron
output:
  html_notebook:
    code_folding: hide
    toc: yes
    toc_depth: 4
    toc_float: yes
runtime: shiny
---

```{r setup, include=FALSE}
library(tidyverse)
library(data.world)
library(shiny)
library(fiftystater)
library(mapproj)

knitr::opts_chunk$set(echo = TRUE)
```
  
## **R Session Info**  

```{r}
sessionInfo()
```

## **Github Link** 
https://github.com/devtl/f17-edv-project3

## **Introduction** 
Fall 2017 Data Visualization Project. We want to explore tobacco usage among highschool teenagers using data from the U.S. Census Bureau.

Dataset is stored on https://data.world/devtl/f-17-edv-project-3.

## **Getting data from data.world** 
We will be using the Fatal Police Shootings data and some census data from the U.S. Census Bureau from [data.world](https://data.world/devtl/fatalpoliceshootings/)

```{r}
project <- "https://data.world/devtl/f-17-edv-project-3/"
data.world::set_config(cfg_env("DW_API"))
df <- data.world::query(data.world::qry_sql("select smoking.surveyyear as year,
    smoking.statecode as state,
    poverty.geoid,
    ((poverty.b17001_008 + poverty.b17001_009) + (poverty.b17001_022 + poverty.b17001_023))/poverty.b17001_002 as belowpoverty_highschool,
    (poverty.b17001_008 + poverty.b17001_009)/poverty.b17001_002 as belowpoverty_highschool_m,
    (poverty.b17001_022 + poverty.b17001_023)/poverty.b17001_002 as belowpoverty_highschool_f,
    smoking.EverTryCigT as ever_tried,
    smoking.EverTryCigF as evertried_f,
    smoking.EverTryCigM as evertried_m,
    smoking.CigBefore13T as before13,
    smoking.CigBefore13F as before13_f,
    smoking.CigBefore13M as before13_m,
    smoking.CurrentCigT as currentsmoker,
    smoking.CurrentCigF as currentsmoker_f,
    smoking.CurrentCigM as currentsmoker_m,
    smoking.FrequentCigT as freqsmoker,
    smoking.FrequentCigF as freqsmoker_f,
    smoking.FrequentCigM as freqsmoker_m,
    smoking.TenCigDayT as 10aday,
    smoking.TenCigDayF as 10aday_f,
    smoking.TenCigDayM as 10aday_m,
    smoking.NotTryQuitT as notquit,
    smoking.NotTryQuitF as notquit_f,
    smoking.NotTryQuitM as notquit_m,
    smoking.CigAtSchoolT as atschool,
    smoking.CigAtSchoolF as atschool_f,
    smoking.CigAtSchoolM as atschool_m,
    smoking.BuyCigAtStoreT as buyscigs,
    smoking.BuyCigAtStoreF as buyscigs_f,
    smoking.BuyCigAtStoreM as buyscigs_m,
    smoking.EverCigDailyT as atleast1_daily,
    smoking.EverCigDailyF as atleast1_daily_f,
    smoking.EverCigDailyM as atleast1_daily_m,
    smoking.CigAll30DaysT as 30days,
    smoking.CigAll30DaysT as 30days_f,
    smoking.CigAll30DaysT as 30days_m,
    smoking.CurrentSmokelessT as smokeless,
    smoking.CurrentSmokelessT as smokeless_f,
    smoking.CurrentSmokelessT as smokeless_m,
    smoking.CurrentCigarT as cigar,
    smoking.CurrentCigarF as cigar_f,
    smoking.CurrentCigarM as cigar_m,
    smoking.CurrentAnyTobaccoT as anytobacco,
    smoking.CurrentAnyTobaccoF as anytobacco_f,
    smoking.CurrentAnyTobaccoM as anytobacco_m
    
from health.`yrbs-state-tobacco-variables`.yrbs_state_tobacco_variables_2013_v2 as smoking
join uscensusbureau.`acs-2015-5-e-poverty`.usa_all_states as poverty 
    on smoking.statecode = poverty.state"), dataset = project)
summary(df)
```


##Scatterplots{.tabset .tabset-fade}

Various Tobacco Usages vs. Highschool Poverty Levels

###Any type of Tobacco Usage
```{r}
df1 = df %>% dplyr::select(belowpoverty_highschool, anytobacco, atleast1_daily, atschool, before13, currentsmoker, freqsmoker, state)

renderPlot({
  pl1 = ggplot(df1, mapping = aes(belowpoverty_highschool, anytobacco, color = state)) + geom_point(size = 1.5) + theme_minimal()
  
 pl1 
})
```

###Daily Smokers
```{r}
renderPlot({
  pl2 = ggplot(df1, mapping = aes(belowpoverty_highschool, atleast1_daily, color = state)) + geom_point(size = 1.5) + theme_minimal()
  
  pl2
  
})
```


###Smokes at School
```{r}
renderPlot({
  pl3 = ggplot(df1, mapping = aes(belowpoverty_highschool, atschool, color = state)) + geom_point(size = 1.5) + theme_minimal()
  
  pl3 
})
```


###Smoked before 13 years old
```{r}
renderPlot({
  pl4 = ggplot(df1, mapping = aes(belowpoverty_highschool, before13, color = state)) + geom_point(size = 1.5) + theme_minimal()
  
  pl4

})
```


###Current Smoker
```{r}
renderPlot({
  pl5 = ggplot(df1, mapping = aes(belowpoverty_highschool, currentsmoker, color = state)) + geom_point(size = 1.5) + theme_minimal()
  
  pl5

})
```

###Frequent Smoker
```{r}
renderPlot({
  pl6 = ggplot(df1, mapping = aes(belowpoverty_highschool, freqsmoker, color = state)) + geom_point(size = 1.5) + theme_minimal()
  
  pl6
})
```

###Findings

We see that there is no evidence that poverty affects any smoking factors among highschool teens.


## Map Plot{.tabset .tabset-fade}

###State Poverty Levels

The following maps the percentage of highshool-aged teens who are living in poverty:

```{r}
mapdf <- fifty_states
pov <- aggregate(belowpoverty_highschool ~ state, data = df, FUN = mean)

rep <- data.frame(
  state = c(
    "AL","AK","AZ","AR","CA","CO","CT","DE","FL","GA","HI","ID","IL","IN","IA",
    "KS","KY","LA","ME","MD","MA","MI","MN","MS","MO","MT","NE","NV","NH","NJ",
    "NM","NY","NC","ND","OH","OK","OR","PA","RI","SC","SD","TN","TX","UT","VT",
    "VA","WA","WV","WI","WY"), 
  id = c("alabama", "alaska", "arizona", "arkansas", "california", "colorado", "connecticut", 
         "delaware", "florida", "georgia", "hawaii", "idaho", "illinois", "indiana", "iowa",
         "kansas", "kentucky", "louisiana", "maine", "maryland", "massachusetts", "michigan",
         "minnesota", "mississippi", "missouri", "montana", "nebraska", "nevada", "new hampshire", 
         "new jersey","new mexico", "new york", "north carolina", "north dakota", "ohio", 
         "oklahoma", "oregon", "pennsylvania", "rhode island", "south carolina", 
         "south dakota", "tennessee", "texas", "utah", "vermont", "virginia", "washington", 
         "west virginia", "wisconsin", "wyoming"))

pov <- merge(pov,rep, by = "state")

mapdf1 <- merge(mapdf, pov, by = "id", all.x = TRUE)

# map_id creates the aesthetic mapping to the state name column in your data
renderPlot({
  ggplot(mapdf1, aes(long, lat, map_id = id)) + 
    geom_map(aes(fill = belowpoverty_highschool), map = fifty_states) +
    scale_fill_gradient(low = "light grey", high = "dark blue") +
    coord_map() + labs(fill = "Percent Below Poverty") +
    scale_x_continuous(breaks = NULL, labels = NULL, name = NULL) + 
    scale_y_continuous(breaks = NULL, labels = NULL, name = NULL) +
    theme_minimal()
  
})

```

###Current Smokers

```{r}
smoke <- aggregate(currentsmoker ~ state, data = df, FUN = mean)

smoke <- merge(smoke,rep, by = "state")

mapdf2 <- merge(mapdf, smoke, by = "id", all.x = TRUE)

# map_id creates the aesthetic mapping to the state name column in your data
renderPlot({
  ggplot(mapdf2, aes(long, lat, map_id = id)) + 
    geom_map(aes(fill = currentsmoker), map = fifty_states) +
    scale_fill_gradient(low = "light grey", high = "tomato3") +
    coord_map() + labs(fill = "Average Number of Smokers") +
    scale_x_continuous(breaks = NULL, labels = NULL, name = NULL) + 
    scale_y_continuous(breaks = NULL, labels = NULL, name = NULL) +
    theme_minimal()
  
})
```

